---
title: "Practical Machine Learning - Course Project"
author: "Anastassiya Zidkova"
date: "23.5.2015"
output: html_document
---

##Abstract

##Introduction

##Enabling parallelization
Before performing any analysis we 

```{r}
library(doParallel);
rCluster <- makePSOCKcluster(4);
registerDoParallel(rCluster);
```


##Cleaning data
Data where cleaned by removing variables containing NA values and unnecessary variables ("timestamp", "X", "user_name", "new_window"). Cleaned data contained 54 variables

```{r}
setwd("~/Documents/Coursera/PracticalMachineLearning/CourseProject")
data <- read.csv("pml-training.csv", head=T, na.strings=c("NA",""))
data2 <- data[,colSums(is.na(data)) == 0] 
removeIndex <- grep("timestamp|X|user_name|new_window", names(data2))
data3 <- data2[, -removeIndex]
```

##Creating training and testing datasets
Data were divided into training and testing datasets by following scheme: 70% of data in training set and 30% of data in testing set.


```{r}
library(caret)
set.seed(17)
trainL <- createDataPartition(y=data3$classe, p=0.7, list=FALSE)
train <- data3[trainL,]
test <- data3[-trainL,]
dim(train); dim(test)
```
Training set contained `r nrow(train)` observations, testing set contained `r nrow(test)` observations.

##Fitting linear regression model

```{r}
set.seed(17)
modFitAll<- train(classe ~ .,data=train,method="rf", trControl = trainControl(method = "cv", number=4))
pred <- predict(modFitAll, test)
modFitAll$results
cfM <- confusionMatrix(pred, test$classe)
cfM$table
```
##Predicting values of validation set

```{r}
set.seed(17)
valid <- read.csv("pml-testing.csv", head=T, na.strings=c("NA",""))
valid2 <- valid[,colnames(valid)%in%colnames(train)] 
pred2 <- predict(modFitAll, valid2)
valid3 <- valid2
valid3$classe <- pred2

pml_write_files = function(x) {
    n = length(x)
    for (i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
            col.names = FALSE)
    }
}


answers <- as.character(valid3$classe)

pml_write_files(answers)
answers

```

##Conclusion

##References
[1] Ugulino, Wallace, et al. "Wearable computing: accelerometers’ data classification of body postures and movements." Advances in Artificial Intelligence-SBIA 2012. Springer Berlin Heidelberg, 2012. 52-61.
